---
title: "round robin simulations"
format: html
editor: visual
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(warning = FALSE, message = FALSE) 
```

# Bradley Terry

teams and strengths

```{r}
set.seed(4329)
library(tidyverse)

#set up
num_teams <- 8
teams <- paste("Team", 1:num_teams)

team_strengths <- data.frame(Team = teams,
  Strength = round(runif(num_teams, min = 0, max = 1),2))

# add column for true ranks of team_strengths
(team_strengths <- team_strengths %>% arrange(-Strength) %>% mutate(Rank = row_number()) %>% arrange(Team))
```

simulate one match using Bradley Terry model
```{r}
# simulate a match using Bradley-Terry model
simulate_match <- function(team1, team2, strengths) {
  strength1 <- strengths$Strength[strengths$Team == team1]
  strength2 <- strengths$Strength[strengths$Team == team2]
  
  prob_team1_wins <- strength1 / (strength1 + strength2)  # BT model probability
  
  winner <- ifelse(runif(1) < prob_team1_wins, team1, team2)  # Randomize outcome
  return(winner)
}
```

simulate single round robin tournament

```{r}
simulate_roundrobin <- function(teams, strengths){
  n <- length(teams)
  win_counts <- numeric(n)
  for (i in seq_along(teams)){
    for (j in seq_along(teams)){
      if (i != j){
        match_winner <- simulate_match(teams[i],teams[j],strengths)
        if (match_winner == teams[i]) {
          win_counts[i] <- win_counts[i] + 1
        } else {
          win_counts[j] <- win_counts[j] + 1
        }
      }
    }
  }
  results <- data.frame(
    Team = teams,
    Wins = win_counts
  )
  
  # Ranks teams based on number of wins (includes ties)
  results$RankHat <- rank(-results$Wins, ties.method = "average")
  results <- results[order(results$RankHat),]
  
  results <- results %>% inner_join(team_strengths, by = 'Team') %>% select(Team, Strength, Rank, RankHat, Wins)
  
  return(results)
}

rr <- simulate_roundrobin(teams, team_strengths)

# kendall correlation
kendallcor <- cor.test(rr$RankHat,rr$Rank, method="kendall")
```

simulate many round robin tournaments using the same set of 8 teams

```{r}
# Define the number of simulations
n <- 1000  # Set the number of simulations you want
total_ranks <- data.frame(matrix(NA, nrow = n, ncol = length(teams)))
colnames(total_ranks) <- teams

kendallcor <- c()

# Run multiple simulations
for (i in 1:n) {
  simulation <- simulate_roundrobin(teams, team_strengths)  # Call the tournament simulation function
  ranks <- simulation$Rank
  team_order <- simulation$Team
  total_ranks[i, team_order] <- ranks
  
  k <- cor.test(simulation$RankHat,simulation$Rank, method="kendall")$estimate
  kendallcor[i] <- k
}

# Calculate average ranks for each team
average_ranks <- data.frame(AvgRankHat = colMeans(total_ranks, na.rm = TRUE), Strength = team_strengths$Strength, Rank = team_strengths$Rank)
average_ranks

# distribution of kendall coefficients
hist(kendallcor)
mean(kendallcor)
median(kendallcor) # rank hat matches true rank about 70% of the time
```

# Finalizing Round Robin Function

### generate winning probabilities with$\exp(i-j)/(1+\exp(i-j))$

round robin function inputs: teams, strengths, schedule, (seed?)

## create teams and strengths - strengths are just 1-8

```{r}
library(tidyverse)

# Define the number of teams 
num_teams <- 8
teams <- paste("Team", 1:num_teams)

# Assign random strengths to each team 
team_strengths <- data.frame(
  Team = teams,
  Strength = (1:num_teams)
)

# add column for true ranks of team_strengths
team_strengths <- team_strengths %>% arrange(-Strength) %>% mutate(Rank = row_number()) %>% arrange(Team)
```

## simulate a single match

```{r}
simulate_match <- function(team1, team2, strengths) {
  strength1 <- strengths$Strength[strengths$Team == team1]
  strength2 <- strengths$Strength[strengths$Team == team2]
  
  prob_team1_wins <- exp(strength1-strength2) / (1 + exp(strength1-strength2))
  
  winner <- ifelse(runif(1) < prob_team1_wins, team1, team2)
  return(winner)
}
```

## simulate a single round robin tournament

```{r}
simulate_roundrobin <- function(teams, strengths){
  n <- length(teams)
  win_counts <- numeric(n)
  for (i in seq_along(teams)){
    for (j in seq_along(teams)){
      if (i != j){
        match_winner <- simulate_match(teams[i],teams[j],strengths)
        if (match_winner == teams[i]) {
          win_counts[i] <- win_counts[i] + 1
        } else {
          win_counts[j] <- win_counts[j] + 1
        }
      }
    }
  }
  results <- data.frame(
    Team = teams,
    Wins = win_counts
  )
  
  # Ranks teams based on number of wins (includes ties)
  results$RankHat <- rank(-results$Wins, ties.method = "average")
  results <- results[order(results$RankHat),]
  
  results <- results %>% inner_join(team_strengths, by = 'Team') %>% select(Team, Strength, Rank, RankHat, Wins)
  
  return(results)
}

rr <- simulate_roundrobin(teams, team_strengths)
# kendall correlation
kendallcor <- cor(rr$RankHat, rr$Rank, method="kendall")
# spearman correlation
spearmancor <- cor(rr$RankHat, rr$Rank, method="spearman")
```

Kendall Coefficient

Kendall Coefficient is a statistic used to measure the ordinal association between two measured quantities.

-   If the agreement between the two rankings is perfect (i.e., the two rankings are the same) the coefficient has value 1.
-   If the disagreement between the two rankings is perfect (i.e., one ranking is the reverse of the other) the coefficient has value −1
    -   We do not have any values of the coefficient that are less than 0

Spearman Coefficent

Spearman Correlation is a non-parametric correlation also known as rank-based correlation coefficient. It measures the strength and direction of the monotonic relationship between two ranked variables. It ranges from -1 to +1:

-   **+1**: A perfect positive monotonic relationship.

-   **0**: No monotonic relationship.

-   **-1**: A perfect negative monotonic relationship.

### simulate many round robin tournaments using the same set of 8 teams, get kendall and spearman coefficients

```{r}
# Define the number of simulations
n <- 1000  # Set the number of simulations
total_ranks <- data.frame(matrix(NA, nrow = n, ncol = length(teams)))
colnames(total_ranks) <- teams

kendallcor <- c()
spearmancor <- c()

# Run multiple simulations
for (i in 1:n) {
  simulation <- simulate_roundrobin(teams, team_strengths)  # Call the tournament simulation function
  ranks <- simulation$Rank
  team_order <- simulation$Team
  total_ranks[i, team_order] <- ranks
  
  kendallcor[i] <- cor(simulation$RankHat, simulation$Rank, method="kendall")
  spearmancor[i] <- cor(simulation$RankHat, simulation$Rank, method="spearman")
}

# Calculate average ranks for each team
average_ranks <- data.frame(AvgRankHat = colMeans(total_ranks, na.rm = TRUE), Strength = team_strengths$Strength, Rank = team_strengths$Rank)
average_ranks

# distribution of kendall correlation
print('kendall')
hist(kendallcor)
mean(kendallcor)
median(kendallcor)

# distribution of spearman correlation
print('spearman')
hist(spearmancor)
mean(spearmancor)
median(spearmancor)
```






# ROUND ROBIN CODE

## simulate a single match

```{r}
simulate_match <- function(team1, team2, strengths) {
  strength1 <- strengths$Strength[strengths$Team == team1]
  strength2 <- strengths$Strength[strengths$Team == team2]
  
  prob_team1_wins <- exp(strength1-strength2) / (1 + exp(strength1-strength2))
  
  winner <- ifelse(runif(1) < prob_team1_wins, team1, team2)
  return(winner)
}
```

## simulate a single round robin tournament

```{r}
simulate_roundrobin <- function(teams, strengths){
  n <- length(teams)
  win_counts <- numeric(n)
  for (i in seq_along(teams)){
    for (j in seq_along(teams)){
      if (i != j){
        match_winner <- simulate_match(teams[i],teams[j],strengths)
        if (match_winner == teams[i]) {
          win_counts[i] <- win_counts[i] + 1
        } else {
          win_counts[j] <- win_counts[j] + 1
        }
      }
    }
  }
  results <- data.frame(
    Team = teams,
    Wins = win_counts
  )
  
  # Ranks teams based on number of wins (includes ties)
  results$RankHat <- rank(-results$Wins, ties.method = "average")
  results <- results[order(results$RankHat),]
  
  results <- results %>% inner_join(team_strengths, by = 'Team') %>% select(Team, Strength, Rank, RankHat, Wins)
  
  return(results)
}
```


# code that is changed for different team strengths/number sims, etc.
## generate team strengths using normal distribution
```{r}
# Define the number of teams 
num_teams <- 8
teams <- paste("Team", 1:num_teams)

# Assign random strengths to each team 
team_strengths <- data.frame(
  Team = teams,
  Strength = rep(0,8))

# add column for true ranks of team_strengths
team_strengths <- team_strengths %>% arrange(-Strength) %>% mutate(Rank = row_number()) %>% arrange(Team)
team_strengths
```

run round robin simulation
```{r}
# Define the number of simulations
n <- 1000  # Set the number of simulations
total_ranks <- data.frame(matrix(NA, nrow = n, ncol = length(teams)))
colnames(total_ranks) <- teams

kendallcor <- c()
spearmancor <- c()

# Run multiple simulations
for (i in 1:n) {
  simulation <- simulate_roundrobin(teams, team_strengths)  # Call the tournament simulation function
  ranks <- simulation$Rank
  team_order <- simulation$Team
  total_ranks[i, team_order] <- ranks
  
  kendallcor[i] <- cor(simulation$RankHat, simulation$Rank, method="kendall")
  spearmancor[i] <- cor(simulation$RankHat, simulation$Rank, method="spearman")
}

# Calculate average ranks for each team
average_ranks <- data.frame(AvgRankHat = colMeans(total_ranks, na.rm = TRUE), Strength = team_strengths$Strength, Rank = team_strengths$Rank)
average_ranks

# distribution of kendall correlation
print('kendall')
hist(kendallcor)
mean(kendallcor)
median(kendallcor)

# distribution of spearman correlation
print('spearman')
hist(spearmancor)
mean(spearmancor)
median(spearmancor)
```

# controlling the strengths but still have normal distribution
```{r}
thetas <- qnorm((seq(0,1,length=9) + 1/16)[-9])
thetas

# qq plot way
qnorm(((1:8) - 0.5) / 8, 0, 0.5)

# plot with variability on x and average of correlation coefficient on y

# qunif
```
