---
title: |
  | \large Assessing Ranking Fidelity of Competition Structures via Weighted Mutual Information
abstract: | 
 | We evaluate multiple competitions structures to assess their effectiveness in accurately reproducing the true underlying ranking of participating teams. Using repeated simulations for each structure, we compare teams’ true performance ranks to their simulated outcomes. To quantify the accuracy of each format, we calculate weighted mutual information between true and simulated rankings across a large number of replicates. "Salient" weights are applied to look at only the top i out of n total teams/competitors in a given competition structure. This information-theoretic approach allows us to objectively compare formats and identify which structures most effectively preserve rank order. Results are visualized through comparative plots, providing clear insights into the trade-offs and strengths of each competition design.
 | 
  \vspace{2mm}
  | Keywords: tournament structures, mutual information
bibliography: references.bib
fontsize: 12pt
link-citations: true
linkcolor: cyan
urlcolor: cyan
output:
  pdf_document:
    df_print: kable
    number_sections: true
    keep_tex: true
header-includes:
 \usepackage{setspace}
 \setstretch{1.15}
 \usepackage{float}
 \floatplacement{figure}{t}
editor_options: 
  markdown: 
    wrap: 72
---

```{r setup, echo=FALSE, message=FALSE}
knitr::opts_chunk$set(
	echo = FALSE,
	fig.align = "center",
	message = FALSE,
	warning = FALSE
)

gjm <- function(x, color = "red") {
  sprintf("\\textcolor{%s}{%s}", color, x)}
```

```{r pkgs}
library(tidyverse)
library(dplyr)
library(stringr)
theme_set(theme_minimal())
```

# Introduction

<!-- @Rajski1961 -->

<!-- @McShane2019 -->

<!-- Csato 2020: The incentive (in)compatibility of group-based qualification systems. World cup qualifying has problems where teams are better off exerting lower effort -->

<!-- @Sziklai2022 -->

There are hundreds of different competitions structures that dictate how
a set of teams compete against each other to determine an overall
ranking. Common structures include bracket tournaments where if a team
wins their competition they advance to the next round. Bracket
tournaments can be run as single elimination where competitors advance
or are eliminated based on a single match (e.g. NCAA College Basketball
Tournament, National Football League (NFL) playoffs, major tennis
tournaments) or where each round is contested as a best of $k$ series
(e.g. $k$ = 7: National Basketball Association (NBA) playoffs, National
Hockey League (NHL) playoffs, $k$ varies by round: Major League Baseball
(MLB), Women's National Basketball Association (WNBA)) and the series
winner advance while the series loser is eliminated. Additional styles
of the bracket tournaments include variations where competitors are not
eliminated after a single loss such as double elimination tournaments
where competitors continue until they have lost twice. Within double
elimination tournament, there are several variations. For instance, the
NCAA College Wrestling tournament is a double elimination tournament,
however, if a competitor loses a match they fall into the consolation
bracket and the best they can finish after a loss before the finals is
3rd place making this not a "true" double elimination format. Olympic
judo uses a variant of this system with single elimination until the
quarterfinals followed by a consolation bracket consisting of the four
quarterfinal losers who still have a chance for a bronze medal[^1].

[^1]: in Olympic judo they award two bronze medals to the two
    competitors who each win their side of the consolation bracket

"True" double elimination - a bracket tournament where a competitor can
still finish ranked first as long as a competitor has not lost two
games - is used in NCAA baseball and softball for certain rounds in
their postseason. Additional variants of double elimination include
double elimination brackets with a "true third" match which includes an
additional match contested between the competitors ranked 2nd (loser of
the finals) and 3rd (winner of the consolation bracket) at the end of
the tournament *if they have not already competed against each other*.
There is also a variation of double elimination where only those
competitors who have lost to a competitor that makes the finals are
placed into the consolation bracket and given the chance to compete for
third place. This style of competition is used in Olympic wrestling
(both freestyle and Greco-Roman)[^2].

[^2]: As in Olympic judo, Olympic wrestling also award two bronze medals
    to the winners of each side of the consolation bracket

<!-- A paragraph about Swiss, Dutch, and Monrad systems -->

In bracket style competitions, only a handful of the match-ups that are
possible actually occur. When every possible competitive pairing occurs,
this is referred to as a round-robin structure where every competitor
plays every other competitor the same number of times. Single
round-robin, where was possible match-up occurs exactly once, is used
in, for examples, the group stage of the FIFA World Cup to create a
ranking to determine which teams advance to the next round. Double
round-robin, where each possible pairing occurs twice, is used in
association football in English Premier League (EPL) and in chess during
the FIDE candidates tournament (a tournament that decides who will
challenge the world champion). For the sake of brevity, we review only a
small number of possible competition structures, but the interested
reader can find a more complete list of competition structures in
@Goosens2024, @Appleton1995-zn and @Csato2021.

@Daniels1969

A natural question of interest after reviewing different types of
competition structures is: which competition structure is best? As with
most complicated questions, the answer is "it depends". One commonsense
conceptualization for the best competition structure is that the final
rankings produced by a competition structure should closely match the
true rankings of the strength of the competitors. That is, a good
competition structure should rank the true best team first, the true
second best team second, and so on, until finally ranking the true worst
team last. We refer to this property of a competition structure as the
efficacy of the structure (@Lasek2018-de, @Sziklai2022-kf). If one were
only interested in correctly ranking the true best competitor as first,
with the ranks of the remaining competitors viewed as inconsequential,
we refer to this here are effectivity (@Glenn1960-mn). While there are
many competition structures where efficacy and effectivity are useful
measures[^3], there are many other competitions where the correct
ranking of more than one competitor, but not all the competitors is of
import. For instance, in Olympic competition, the goal of a competition
structure is to correctly rank the top 3 competitors to award gold,
silver, and bronze medals to[^4]. In many American sports, for instance
the NFL, the goal of the competition structure of the regular season is
to rank the top 7 teams in each conference to choose the playoff teams
with the ranking of the teams 8 through 16 in each conference
inconsequential in determining a Super Bowl champion[^5]. We coin the
term "salient ranks" to refer to the ranks of a competition structure
that matter. That is the ranks 1, 2, and 3 in Olympic competition are
the salient ranks in that competition structure. In this notation,
efficacy is a measure with with salient ranks 1 through $n$ where $n$ is
the number of competitors, and effectivity is a measure where 1 is the
only salient rank.\
<!-- Efficacy: @Lasek2018-de, @Sziklai2022-kf -->
<!-- Effectivity: @Glenn1960-mn -->

[^3]: efficacy is a useful measure for the EPL, for example, as the
    ranking of the teams matters at both the high end (i.e. entry into
    special tournaments) and the low end (i.e. relegation status) and
    effectivity is a useful measure for the NCAA College Basketball
    tournament (i.e. main goal is to crown a champion)

[^4]: Some sports award two bronze medals (e.g. wrestling, judo, boxing,
    tae kwon do)

[^5]: The ranking of teams outside of the playoffs does matter for draft
    purposes though

<!-- From Goosens 2024: The ability of a tournament to rank contestants according to their true strength is called efficacy (Lasek and Gagolewski, -->

<!-- 2018; Sziklai et al., 2022). Efficacy with respect to only the best player is also known as effectivity (Glenn, 1960) -->

<!-- or predictive power (Ryvkin and Ortmann, 2008; Vu and Shoham, 2011). -->

If the only goal of a competition structure was high efficacy of the
salient ranks, the best way to achieve this would be to play each pair
of possible match ups of competitors as many times as possible. Of
course, this is impossible due to many factors including time
constraints, cost constraints, and wear and tear on competitors in high
impact sports (e.g. boxing, American football, etc.). In addition,
competitions where the true best team is guaranteed or nearly guaranteed
to win may be less appealing to fans and may drive down demand as
discussed in @Johnson2022-wz.

In light of this, many different competition structures have been
proposed, several of which are mentioned previously. Given these many
competition structures, ideally there would be a way comparing the
performance of different competition structures to each other.

Here, we propose a novel metric for evaluating different competition
structures using an information theoretic approach. As an analogy, we
view the results of a tournament as a transmitted message, and we use
information theory to assess how accurately the message (i.e. the
tournament results) matches the true transmission (i.e. the true ranking
of competitors).

Specifically, we use weighted mutual information with a weights based on
squared error loss[^6] between the ranks generated by a competition
structure and the true ranks of the competitors. Additionally, we
include an indicator vector to specify the salient ranks, allowing us to
compare different competition structures with different choices for
salient ranks.

[^6]: other loss functions can easily be incorporated into our method

The remainder of this manuscript presents the mathematical foundation of
the proposed measure followed by a simulation study estimating out
proposed measure for different types of competition structures with
varying choices of salient ranks. We close by offering a summary of the
work and ideas for future directions.

<!-- @Marchand2002-tq only cares about the probability of winning in seeded vs random draw knockout tournemnts.   -->

<!-- @Johnson2022-wz: Match outcome uncertainty and the Rottenberg hypothesis -->

<!-- Different competition structures have unique strengths and weaknesses, affecting how well they reflect the true rankings of the teams. The goal of a tournament is to find the best teams, but how can we quantify how well a tournament performs? In some instances, the tournament organizers only award the overall winner, but other times, the top three or ten teams are awarded. Because of this, organizers may prefer one structure over another based on their needs. Tournament organizers must also take into account factors like cost, timeliness, entertainment value, and fairness when choosing a tournament structure. -->

<!-- To evaluate a tournament’s effectiveness, we propose a numerical metric that quantifies how accurately it orders teams based on their true rankings. Tournament results can be viewed as a "message" attempting to convey the true team rankings. However, various factors introduce noise, leading to information loss. We use principles from information theory to measure this information loss across multiple tournament simulations to assess the reliability of different formats. -->

# Methods

Consider a set of $n$ teams indexed from $i = 1, 2, \cdots, n$ each with
an associated true strength parameter
$\boldsymbol{\theta} = (\theta_1, \theta_2, \ldots, \theta_n)$ such that
$(\theta_1 \ge \theta_2 \ge \ldots \ge \theta_n)$. Next, let
$r(\boldsymbol{\theta})$ be the indexes of $\boldsymbol{\theta}$ when
$\theta$ is sorted from largest to smallest so that
$r(\boldsymbol{\theta}) = \{1, 2, \ldots, n-1, n\}$ (Note: While we
allow for two values of $\theta$ to be equal to each other, we choose to
assign unique ranks to each $\theta_i$). Next define $T(\phi, s)$ as a
tournament with schedule $\phi$ and seeding structure $s$. $\phi$
contains all the information about the scheduling of teams, which could
be fully known prior to the tournament (*e.g.* round robin) or
determined as the tournament progresses (*e.g.* single elimination
tournament). We then let $\hat{r}(T(\phi, s),\boldsymbol{\theta})$ be an
$n$-dimensional vector-valued random variable that gives the results of
a tournament as a vector of the indexes of $\boldsymbol{\theta}$. The
number in the first position of the vector
$\hat{r}(T(\phi, s),\boldsymbol{\theta})$ indicates the index of the
team that finished first, the number in the second position of the
vector $\hat{r}(T(\phi, s),\boldsymbol{\theta})$ indicates the index of
the team that finished second, and so on. (Note: We do allow for ties in
$\hat{r}$.)

For example, in a $n$ = 4 team tournament,
$\hat{r}(T(\phi, s),\boldsymbol{\theta})$ = (3, 2, 1, 4) indicates that
the team with index 3 (i.e. the true third best team) won the
tournament, the true second best team finished second, the true best
team finished third and the true 4th team finished 4th in that
particular tournament. If $\hat{r}(T(\phi, s),\boldsymbol{\theta})$ =
(1, 2, 3, 4), this means that the random outcome of the tournament was
the same as the true ordering of the teams.

As noted above we allow for ties in $\hat{r}$. Continuing the above
example, if $\hat{r}(T(\phi, s),\boldsymbol{\theta})$ = (3, 2, 1, 3)
this indicates that the teams that are truly ranked 1st and 4th tied for
3rd.

We need a transition: If one views the true ranking of the teams in a
tournament as a message to be sent to a receiver and the outcome of the
tournament as a message that is received, we can measure the "goodness"
of a tournament in terms of its ability to accurately transmit the true
ranking. We can then use concepts from information theory to assess the
ability of a tournament to correctly rank teams. Specifically, we start
with the concept of mutual information (@Guiasu1977-nu).

For two random variables $X$ and $Y$ mutual information is defined as:

$$
I(X;Y)=\sum_{y\in Y}\sum_{x\in X}p(x,y)\log \frac {p(x,y)}{p(x)\,p(y)}
$$

In our setting, we replace $X$ and $Y$ with $r(\boldsymbol{\theta})$ and
$\hat{r}(T(\phi, s),\boldsymbol{\theta})$, and we could seek to estimate
$I(r(\boldsymbol{\theta}),\hat{r}(T(\phi, s),\boldsymbol{\theta}))$. For
simplicity, we drop the arguments from the functions $r$ and $\hat{r}$
for ease of exposition.

In order to compute mutual information, marginal distributions of $r$
and $\hat{r}$ are required. Without knowing $r$, all possible
permutations of $\hat{r}$ are equally likely so we set
$p(\hat{r}) = \frac{1}{n!}$. By a similar argument we set
$p(r) = \frac{1}{n!}$.

we set . By a similar argument we set $p(\hat{r}) = \frac{1}{n!}$.

We define the mutual information of $r$ and $\hat{r}$ to be:

$$
I(r,\hat{r})=\sum_{r}\sum_{\hat{r}}p(r,\hat{r})\log \frac {p(r,\hat{r})}{p(r)p(\hat{r})} = n!\sum_{\hat{r}}p(r,\hat{r})\log \frac {p(r,\hat{r})}{(\frac{1}{n!})^2}
$$

While there are $n!$ different permutations for the result of $r$, we
don't need to sum across these as any specific choice of $r$ is just an
arbitrary labeling of the true strength parameters vector
$\boldsymbol{\theta}$. So given $r$, the distribution of $\hat{r}$ is
the same up to the labeling. Therefore, we only need to consider a
single permutation of $r$, we compute the the quantity inside the
summation, and then multiple by $n!$ (effectively summing across $r$).

In order to compute this quantity, we need to compute $p(r,\hat{r})$.
This probability is found by calculating the probability of a given
permutation of $\hat{r}$ and $r$ is assumed to be the permutation from
$\{1, 2, \ldots, n\}$. We estimate $p(r,\hat{r})$ empirically through
simulation.

However, using mutual information in this form does not suit our needs
in this setting. The problem is that mutual information in this form
will yield high values of mutual information when the output from the
tournament $\hat{r}$ is highly consistent *even if the ranking from the
tournament is incorrect* (*i.e.* the rankings produced by the tournament
are very different from the true rankings). As an example, if $n=4$ and
the true order of $\boldsymbol{\theta}$ is $r = \{1, 2, 3, 4\}$ and
$p(\hat{r} = \{4, 3, 2, 1\}) = 1$ will be the same mutual information as
when $r = \{1, 2, 3, 4\}$ and $p(\hat{r} = \{1, 2, 3, 4\}) = 1$ and in
both of these cases the mutual information will be maximized at:

$$
I(r,\hat{r})= n!\sum_{\hat{r}}p(r,\hat{r})\log \frac {p(r,\hat{r})}{\left(\frac{1}{n!}\right)^2} = n! log((n!)^2)
$$ and for the specific case when $n=4$ would be: $$
4! log_2(4!^2) = 24*log_2(24^2) = 220.0782
$$

# Weighted Mutual Information

In order to alleviate this problem, we instead consider *weighted*
mutual information (@Guiasu1977-nu). We want to give more weight to
permutations from $\hat{r}$ that are "closer" to $r$. Essentially, this
problem comes down to choosing a metric that measures the closeness
between the vectors $\hat{r}$ and $r$. One then gives more weight to
combinations of $\hat{r}$ and $r$ that are closer and less weight to
combinations that are far apart. A natural choice for distance between
vectors the sum of squared deviations between the two vectors.
Specifically, we define the loss function between $\hat{r}$ and $r$ to
be $l(r, \hat{r}) = \sum_{i=1}^n (\hat{r}_i - r_i)^2$. Then, since small
losses should be associated with large weights, we focus on the
recipricol of this loss function.

In general, any loss function $l(r, \hat{r})$ can be used to define the
weighting function as follows:

$$
w(r, \hat{r}) =
\begin{cases}
\frac{1}{l(r, \hat{r})}, & r \ne \hat{r} \\
1, & r = \hat{r}
\end{cases}
$$

Note that in order to avoid dividing by 0, we define the weight to be 1
when $r$ and $\hat{r}$ are exactly equal.

Specifically, when squared error loss is chosen, one gets the following
weight function:

$$
w(r, \hat{r}) =
\begin{cases}
\frac{1}{\sum_{i=1}^n (\hat{r}_i - r_i)^2}, & r \ne \hat{r} \\
1, & r = \hat{r}
\end{cases}
$$ <!-- Note that one is 1/N! -->

<!-- And this does work because consistently getting the order incorrect is perfect information. -->

An alternative to defining weight this way is to consider looking at the
correlation between $\hat{r}$ and $r$. Since, $\hat{r}$ and $r$ are
already ranks, the Pearson and Spearman correlation coefficients will be
computationally exactly the same in this setting if that was the measure
that someone chose to use.

Spearman correlation can be written as follows:

$$
\rho_{s} =  1 - \frac{6  \sum_{i=1}^n (\hat{r}_i - r_i)^2}{n(n^2 - 1)}
$$,

which means that Spearman correlation is simply (one minus) a scaling of
squared error loss. Since $\rho_{s} \in [-1,1]$, but we desire weights
$w(\hat{r},r) \in [0,1]$, we define
$w(\hat{r},r) = \frac{(\rho_s + 1)}{2}$ guaranteeing weights in the
desired range.

While squared error loss and Spearman/Pearson correlation work well, in
this setting we argue that Kendall's $\tau$ correlation is the most
appropriate for weighting due to its interpretation and how this related
to competitions. One way of interpreting Kendall's $\tau$ is that it's a
scaling of the number of adjacent swaps required to go from one
permutation of ranks to another. Recall that Kendall's $\tau$ is defined
as: $$
\rho_\tau = 1 - \frac{2 \times \# \text{ discordant pairs}}{{n \choose 2}}
$$. To convert $\rho_\tau$ back to the range between $[0,1]$, which we
want for weighting purposes, we subtract 1 and divide by 2 giving us
$\frac{\rho_\tau + 1}{2}$. This then becomes:

$$
w(\hat{r},r) = \frac{\rho_\tau + 1}{2} = 1 - \frac{\# \text{ discordant pairs}}{{n \choose 2}}
$$ and the number of discordant pairs is equivalent to the number of
adjacent swaps needed to get the order of $\hat{r}$ correct.\
As an example, consider $r = (1,2,3,4)$ and $\hat{r} = (1,3,4,2)$. This
requires two adjacent swaps to get $\hat{r}$ to equal $r$ (i.e swap 4
with 2 and then 3 with 2). The Kendall's $\tau$ value between these two
vectors is $\frac{1}{3}$. We would then define the weight between the
two vectors to be $(\rho_{\tau} + 1)/2 = \frac{2}{3}$ . This is exactly
equivalent to $1 - \frac{\#\text{ adj swaps}}{{n \choose 2}}$.

<!-- `r cor(c(1,2,3,4),c(1,3,4,2), method = "kendall")` -->

<!-- `r (cor(c(1,2,3,4),c(1,3,4,2), method = "kendall") + 1)/2` -->

<!-- ```{r} -->

<!-- cor(c(1,2,3,4),c(1,3,4,2), method = "kendall") -->

<!-- 1 - (2*2)/choose(4, 2) -->

<!-- cor(c(1,2,3,4),c(1,2,4,3), method = "kendall") -->

<!-- 1 - (2*1)/choose(4, 2) -->

<!-- #Converting this back to 0,1 gives (by adding 1 and dividing by 2):  -->

<!-- #(Number of swaps)/choose(4, 2).   -->

<!-- #This is what we use for weights -->

<!-- ``` -->

Once an appropriate weight function $w(r, \hat{r})$ is chosen, we get
the following formula for weighted information:

$$
I_w(r,\hat{r})= \sum_{r}\sum_{\hat{r}}w(r, \hat{r})p(r,\hat{r})\log \frac {p(r,\hat{r})}{p(r)p(\hat{r})}
$$ \# Tiered Ranking

This framework was developed to assess the efficacy of the full ranking
of the top $k$ teams generated by different competition formats.
However, it may be of interest to have a competition structure that
simply selects the top $k$ teams in any order. Continuing with the
Olympics example, if an organizer only cares about the top 3 teams
earning medals *in any order* (i.e. $\hat{r} = \{1,2,3,4\}$ and
$\hat{r} = \{3,2,1,4\}$ are equivalent), our framework can easily handle
this modifying $\hat{r}$ (and $r$?) so that the top $k$ elements are
ordered correctly and tied. That is, for the Olympic example, if $k$ was
3 and $\hat{r} = \{3,2,1,4\}$, we convert all the elements of
$\hat{r} \le k$ to 1 (i.e. $\hat{r}$ becomes $\{1,1,1,4\}$), and then we
proceed with the exact same formula. We refer to this as tiered ranking.

This concept can be extended to multiple tiers, which is applicable to a
competition structure such as the premier league where the top 4 teams
qualify for the Champions league and the bottom 3 teams are relegated.
While this straight foward to compute, we do not evaluate scenarios such
as this in this article.

# Salient Weights (and ranks)

<!-- @Shieh1998: Weighted Kendall's Tau -->

In addition to the the weighting function $w$, we defined a second set
of weights that we refer to as the "salient weights" and denote as the
vector $m$. This set of weights is used to indicate which ranks have
been determined to salient (as previously described) and consists of a
vector of length $n$ with elements that are positive for each of the
salient ranks and 0 otherwise. As an example, for a four-team
tournament, if the organizer decided that only the correctly choosing
the best team as the winner was important, then $m = \{1,0,0,0\}$ (i.e.
measuring effectivity). In an event such as the Olympics, the salient
weights could be 1's for the first three elements and 0's elsewhere such
that $m = \{1,1,1,0\}$. Additionally, one can also weight the salient
ranks. For example, continuing the Olmypics example where getting the
top three correct is the main objective, one could argue that it is more
important to get the first place correct than the third place. This
could lead to, for instance, a salient weight vector of
$m = \{5,3,1,0, \cdots, 0\}$ giving 5 times the weight to getting the
first place team correct as getting the third team correct.

<!-- Regardless of which method one chooses for defining weights between $\hat{r}$ and $r$, a second set of weights needs to be defined because the number of teams that need to be accurately ordered differs based on the choice of salient ranks. Therefore, we define a second vector of weights, $m$, which is used to indicate which ranks are salient. We call these the "salient weights" to differentiate from the other weighting function, $w$, becomes a function of $m$ (and also, as before, $r$ and $hat{r}$).  As an example, for a four-team tournament, if the organizer decided that only the correctly choosing the best team as the winner was important, then $m = \{1,0,0,0\}$ (i.e. essentially measuring effectivity)^[7].   -->

<!-- ^[7]: One can imagine that in an event like the Olympics, the salient weights would be 1's for the first three elements and 0's elsewhere (i.e. $m = \{1,1,1,0\}$).     -->

With the addition of the salient weighting vector, $m$, the formula for
weighted mutual information becomes:

$$
I_w(r,\hat{r},m)= \sum_{r}\sum_{\hat{r}}w(r, \hat{r}, m)p(r,\hat{r})\log \frac {p(r,\hat{r})}{p(r)p(\hat{r})}
$$ .

Incorporating $m$ into this function is straightforward for many
weighting functions. We use the weighted Kendall's $\tau$ presented in
@Shieh1998 with a slight modification to the sign function.
Specifically, in the special case that $sign(0) = 1$. Then
$w(r, \hat{r}, m)$ becomes:

$$
w(r, \hat{r}, m) = \frac{\sum_{i \ne j} m^{\star}_{ij}\text{sign}_{(0+)}(r_i - r_j)sign_{(0+)}(\hat{r}_i - \hat{r}_j)}{\sum_{i,j}m^{\star}_{i,j}- \sum_i m^{\star}_{i,i}}
$$ where $m^{\star}_{i,j} = m_{min(i,j)}$ and $sign_{(0+)}(.)$ is the
sign of the input with the special case that a 0 maps to 1.

<!-- $$ -->

<!-- wI_m(r_{\{{i|m_i >0}\}}, \hat{r}_{\{{i|m_i >0}\}})=\sum_{r_{\{{i|m_i >0}\}}}\sum_{\hat{r}_{\{{i|m_i >0}\}}} w_m(\hat{r}_{\{{i|m_i >0}\}}, r_{\{{i|m_i >0}\}}) \times -->

<!-- $$ -->

<!-- $$ -->

<!-- p(\hat{r}_{\{{i|m_i > 0}\}}, r_{\{{i|m_i >0}\}})\times \\  -->

<!-- \log \frac{p(\hat{r}_{\{{i|m_i >0}\}}, r_{\{{i|m_i >0}\}})}{p(r_{\{{i  \mid m_i >0}\}})p(\hat{r}_{\{{i \mid m_i >0}\}})} -->

<!-- $$ -->

Finally, because this mutual information is unitless, the formula can be
standardized to be between 0 and 1 as follows: $$
I^{\star}_w(r,\hat{r},m) = \frac{I_w(r,\hat{r},m)}{H(r, \hat{r})}
$$

where $H(.)$ is the usual Shannon entropy. Using this framework, values
of $I^{\star}_w(r,\hat{r},m)$ near 1 indicated competition structures
with high efficacy. Values near 0.25 indicate tournament structures
where all possible rankings are equally likely and values near 0
indicate rankings that are inefficacious (e.g. complete rank reversal).

Under the assumption of equal probability for all possible permutations
of $\hat{r}$, we can simplify $p(r, \hat{r}) = \frac{1}{n!}$ and
$\sum_r \sum_{\hat{r}} w(r, \hat{r},m) = 0.5 \times n!$. Assuming that
$p(r) = p(\hat{r}) = \frac{1}{n!}$, this results in: $$
I_w(r,\hat{r},m) = \sum_{r}\sum_{\hat{r}}w(r, \hat{r}, m)p(r,\hat{r})\log \frac {p(r,\hat{r})}{p(r)p(\hat{r})} = 0.5 \times n! \times \frac{1}{n!} \times \log(\frac{1/n!}{(1/n!)^2}) = 0.5 \times \log(n!)
$$

<!-- I still need to prove why the sum of weight is 0.5 * n! in theory world -->

Additionally, we can assume that $r$ and $\hat{r}$ are independent and
have equivalent probabilities of $\frac{1}{n!}$. Through the chain rule
detailed in @Cover2006, $$
H(r, \hat{r}) = H(r) + H(\hat{r}|r)
$$ . Assuming independence and equality, $$
H(\hat{r}|r) = H(\hat{r}) = H(r)
$$ . Thus, $$
H(r,\hat{r}) = 2 \times H(r) = - 2 \sum_{r} p(r)\log(p(r)) = -2(n!) \times (\frac{1}{n!}) \times \log (\frac{1}{n!}) = \log((n!)^2)
$$

and the final simplification:

$$
I^{\star}_w(r,\hat{r},m) = \frac{I_w(r,\hat{r},m)}{H(r, \hat{r})} = \frac{0.5\times \log(n!)}{2\times \log(n!)} = \frac{0.5}{2} = 0.25
$$

<!-- Likewise, competition structures that are near 0 by our proposed metric indicate low levels of efficacy. -->

<!-- $$ -->

<!-- max\{H(X), H(Y)\} = \sum_{r}  p(r) \log p(r) = n! \log n! -->

<!-- $$ -->

<!-- # ryan way  -->

<!-- Consider a set of $n$ teams indexed from $i = 1, 2, \cdots, n$ each with an associated true strength parameter $\boldsymbol{\theta} = (\theta_1, \theta_2, \ldots, \theta_n)$ such that $(\theta_1 < \theta_2 < \ldots < \theta_n)$.  Next, let $r(\boldsymbol{\theta})$ be the rank of the vector $\boldsymbol{\theta}$ from largest to smallest so that $r(\boldsymbol{\theta}) = \{n, n-1, \ldots, 2, 1\}$.  Next define $T(\phi, s)$ as a tournament with schedule $\phi$ and seeding structure $s$.  $\phi$ contains all the information about the scheduling of teams, which could be fully known prior to the tournament (i.e. round robin) or determined as the tournament progresses (i.e. single elimination tournament).  We then let  $\hat{r}(T(\phi, s),\boldsymbol{\theta})$ be a vector valued random variable that gives the results of a tournament as a vector of ranks.   -->

<!-- Figure out what $\hat{r}$ should look like.   -->

<!-- Let $\boldsymbol{x} = (x_1, x_2, \ldots, x_n)$ be the permutation of indexes of $\boldsymbol{\theta}$ such that $x_1 = {i | \theta_i = max(\boldsymbol{\theta})}$ -->

# Results

We evaluated tournament structures with n = 16 competitors to
demonstrate the utility of our proposed weighted mutual information
metric. Game outcomes were simulated probabilistically using a logistic
model where the probability that competitor j defeats competitor k is
given by:\
$$
p_{jk} = \frac{e^{\theta_j - \theta_k}}{1 + e^{\theta_j - \theta_k}}
$$ where $p_{jk}$ is the probability that competitor $j$ beats
competitor $k$ and $\theta_j$ and $\theta_k$ are the strength parameters
of teams $j$ and $k$, respectively.

We focus on several simple competition structures focused mainly on
variations of bracket tournaments to demonstrate the proof of concept of
our proposed metric. Each competition structure was simulated 10,000
times to estimate $wI_m(r,\hat{r})$. In all bracket structures, seeding
was assumed to be optimal, reflecting each team's true rank. For teams
eliminated before the finals, ranks were assigned randomly within their
possible range (e.g., semifinal losers were assigned rank 3 or 4 with
equal probability).

<!-- we should include links to the images to the ideal seeding structure that Ryan was working on -->

We examined the following structures with the assumption of the ideal
seeding for each competition:

-   Single Elimination, seeding (1 vs 16, 8 vs 9, 4 vs 13, 5 vs 12, 3 vs
    14, 6 vs 11, 7 vs 10, 2 vs 15)

-   Double Elimination, seeding (1 vs 16, 8 vs 9, 4 vs 13, 5 vs 12, 3 vs
    14, 6 vs 11, 7 vs 10, 2 vs 15)

-   Round Robin

-   Group Stage, 4 groups of 4, seeding (pot 1: 1-4, pot 2: 5-8, pot 3:
    9-12, and pot 4: 13-16)

-   Staged Round Robin, 4 groups of 4, seeding (pot 1: 1-4, pot 2: 5-8,
    pot 3: 9-12, and pot 4: 13-16)

-   Tennis Ladder, seeding (16 vs 15, 16 vs 15 winner vs 14, etc.)

![Weighted mutual information of various competition structures as a
function of the top K teams for three distributions of strengths (Coin
Flip, Exponential, Normal).
\label{fig:fig1}](images/combined_plot.png){width="680"}

<!-- need to include table with the values of weighted mutual information for each plot in the appendix -->

Figure \ref{fig:fig1} presents the normalized weighted mutual
information $wI_m(r,\hat{r})$ as a function of the number of salient
ranks (k) across multiple strength distributions. The tennis ladder
structure consistently achieves the highest $wI_m$ values across nearly
all conditions, outperforming the traditional bracket formats regardless
of k or strength distribution. This superior performance arises from the
ladder's inherent sequential constraint, which limits the extent to
which highly-ranked teams can fall in the final standings $\hat{r}$. The
structure essentially provides a built-in protection mechanism for top
competitors while maintaining competitive integrity. However, this
built-in protection is only applicable for the assumption that the
seeding was done correctly.

Round robin tournaments achieve the highest $wI_m$ value for $k=1$ only
under the exponential distribution, where the increased probability of
top-seed victories (due to larger strength differentials) combines with
the structure's maximal number of matchups to provide the exception to
the tennis ladder structure's superior performance. However, this
advantage disappears under more balanced strength distributions.

## Influence of Strength Distributions

The distribution of the strength parameters, $\theta$, can substantially
influence the competition's information. Under the "Coin Flip"
distribution, where every team has an equal 50% win probability, all
structures exhibit reduced $wI_m$ values compared to scenarios with
heterogeneous strengths. This result is intuitive: when true skill
differences are absent, tournament structures cannot extract meaningful
ranking information beyond random assignment.

The exponential distribution, characterized by larger gaps between top
competitors and smaller gaps among lower-ranked teams, produces distinct
patterns. The disparity between single and double elimination widens
considerably for small k, as the heightened probability of top-seed
victories makes the recovery mechanism of double elimination
particularly valuable for preserving information about elite
competitors. Conversely, under normal and uniform distributions, with
more evenly-spaced strength parameters, this advantage diminishes.

It is important to note that while the theoretical value for round robin
for the distribution "Coin Flip" is 0.25, our simulations produced
slightly different estimates due to finite simulation counts and
inherent stochastic variation.

## Tiered Ranking Analysis

![Tiered weighted mutual information of various competition structures
as a function of the top K teams for three distributions of true
strengths (Coin Flip, Exponential, Normal).
\label{fig:fig2}](images/combined_combinations_plot.png){width="680"}

Figure \ref{fig:fig2} examines the mutual information for tiered
rankings of the top $K$ teams. As $K$ approaches $n=16$, $wI_m$
approaches 1 since all teams fall within the "top 16" and ordering
becomes deterministic. The convergence patterns across structures reveal
their relative effectiveness at different ranking granularities.

Most structures exhibit monotonic increases in $wI_m$ as $K$ increases,
reflecting the accumulating information about progressively larger
subsets of the ranking. However, the group stage format (orange)
displays a distinctive non-monotonic pattern characterized by sharp
discontinuities at $K=12$ and $K=16$. These irregularities arise from
the pot-based seeding structure: each group contains exactly one team
from each pot (ranks 1-4, 5-8, 9-12, 13-16). This design creates
clustering effects where teams finishing in specific group positions
span multiple pots.

The sharp increases observed from $K=11$ to $K=12$ and $K=15$ to $K=16$
reflect the resolution of pot-based uncertainties as additional teams
are included in the top-K tier. The plateaus between these transitions
occur because teams eliminated in the same round receive randomly
assigned ranks within their possible range. For instance, the four teams
finishing last in their respective groups receive ranks 13 through 16,
producing no additional information until all four teams are included.
While subdividing the final ranking into multiple tiers (e.g., knockout
qualifiers, third-place finishers, last-place finishers) would eliminate
these discontinuities, we maintained a unified ranking for
cross-structure comparability.

## Effects of Seeding Quality and Competition Variants

Beyond structure type, specific implementation choices significantly
affect tournament informativeness. Figure \ref{fig:fig3} examines
variants of single elimination and ladder structures under different
seeding scenarios and series lengths.

![Mutual information of single elimination variants as a function of the
top K teams for three distributions (Exponential, Normal, Coin Flip) and
three series lengths (1, 3, 7).
\label{fig:fig3}](images/single_elimination_plot.png){width="680"}

<!-- should I include a zoomed in image of the single elimination yellow, red, and green plots? -->

### Seeding Quality

Seeding quality exerts dramatically different effects depending on
structure type. For the step ladder, seeding determines whether the
structure achieves near-optimal performance or falls below random
assignment. Under optimal seeding (top team plays winner of 2 vs 3 vs
... vs 16), the step ladder requires the best team to win only one
match, the second-best two matches, and so forth, yielding exceptionally
high $wI_m$ values. However, under worst-case seeding (1 vs 2, winner vs
3, etc.), the top team must win 15 consecutive matches to achieve the
correct rank, while the second-best team will finish last if the top
team wins the first match. This creates $wI_m$ values below 0.25 for
large $K$ (worse than random assignment) because the structure
systematically inverts the true ranking.

The strength distribution further modulates this effect. Under
exponential distributions where top teams have higher win probabilities,
badly-seeded step ladders show slightly elevated $wI_m$ for small $K$,
as the best team more reliably wins its 15 matches. Under more balanced
distributions, even small $K$ values produce poor results, as upsets
become more probable throughout the extended gauntlet.

Traditional single elimination brackets prove remarkably robust to poor
seeding when reseeding is employed. A common example of a reseeded
single elimination structure is the NFL playoff format. Figure
\ref{fig:fig3} demonstrates minimal divergence between optimal and
worst-case seeding under reseeding protocols, where remaining teams are
rematched each round (highest remaining seed plays lowest, etc.).
Reseeded formats actually slightly outperform standard brackets with
optimal seeding, suggesting that dynamic seeding adjustments provide
value even when initial seeding is correct. This finding indicates that
when feasible, reseeded single elimination represents a more effective
approach than fixed-bracket formats.

### Series Length

Extending individual matchups from single games to best-of-series
formats predictably increases $wI_m$ across all structures and
distributions. Longer series reduce outcome variance, allowing true
strength differences to more reliably determine results. However, the
magnitude of improvement varies by distribution: exponential
distributions show larger gains from extended series than uniform
distributions, as the longer format more reliably separates competitors
with substantial skill gaps.

## Round Robin Performance Under Extended Competition

When time and number of games is not a constraint, round robin is
typically the most common approach to a structure. Examples of this are
found most often in European soccer leagues, like the English Premier
League.

![Mutual information of round robin structures as a function of the top
K teams for three distributions (Exponential, Normal, Uniform) with
varying numbers of complete rounds (1, 2, 4).
\label{fig:fig4}](images/round_robin_plot.png){width="680"}

<!-- do switching to uniform instead of coin flip make sense here? The coin flip would just have all values getting closer to 0.25, but because of the number of simulations it will likely be about 0.4. -->

Figure \ref{fig:fig4} examines round robin tournaments across multiple
complete rounds, where each team plays every other team 1, 2, or 4
times. As the number of rounds increases, $wI_m$ increases across all
distributions, as repeated head-to-head matchups progressively reveal
true strength rankings. Given infinite rounds, the metric would converge
to 1 for any distribution with heterogeneous strengths.

The exponential distribution exhibits wider separation between curves as
$K$ increases, reflecting the varying difficulty of resolving rankings
at different skill levels. Large strength differentials between top
competitors are quickly resolved with high confidence, while similar
bottom-tier competitors require more matchups to differentiate. This
produces expanding gaps in $wI_m$ as additional rounds
disproportionately clarify the difficult-to-resolve lower ranks.

Conversely, the uniform distribution displays more consistent spacing
between round-count curves and only modest variation around $K=8$,
reflecting the equal spacing of strength parameters. Since all pairwise
differences are approximately equivalent, all rank pairs exhibit similar
resolution difficulty, and increasing roounds improves rankings
uniformly across the entire field. As round counts continue to increase
beyond 4, these curves would be expected to flatten progressively.

## Practical Implementation Considerations

While round robin with extensive rounds theoretically provides maximal
information, real-world constraints on season length, athlete health,
and scheduling feasibility necessitate more limited structures. To apply
our metric to specific real-world competitions, we propose three
potential approaches:

1.  **Season-specific simulation**: Estimate strength parameters for
    multiple historical seasons, simulate tournament outcomes for each
    season independently, and aggregate results to compute an overall
    mutual information score
2.  **Averaged metric approach**: Estimate strength parameters for
    multiple historical seasons, compute $wI_m$ separately for each
    season's simulations, and average the resulting metric
3.  **Pooled strength distribution**: Estimate strength parameters
    across multiple seasons, aggregate them by rank position (e.g.,
    average all championship teams' parameters, all second-place teams'
    parameters, etc.) to create a representative strength distribution,
    then simulate tournaments using this composite distribution.

Each approach offers different trade-offs between capturing seasonal
variation and computational tractability, with the optimal choice
depending on the specific application and available historical data.

## Summary

Our weighted mutual information metric successfully quantifies
tournament structure effectiveness in producing accurate rankings. No
single structure universally dominates across all scenarios: tennis
ladders excel under ideal seeding, round robins provide maximal
information given sufficient time, and reseeded single elimination
offers a robust middle ground. The metric reveals that structure choice,
seeding quality, strength distribution, and series length all interact
to determine tournament informativeness. These results provide a
quantitative foundation for evaluating existing tournament designs and
optimizing structures given real-world constraints on time, resources,
and competitive balance objectives.

# Conclusion {#sec:conclusion}

This study introduced a weighted mutual information framework to
evaluate the effectiveness of different competition formats in
preserving the true underlying rankings of teams. We also introduce the
concept of salient ranks, thereby extending and generalizing the
concepts of efficacy and effectivity. By simulating outcomes of
competition structures with varying seeding methods and applying an
information-theoretic approach, we quantified how accurately each format
conveys ranking information. Our findings confirm the intuitive
advantage of round robin tournaments, which consistently outperformed
single and double elimination formats in ranking accuracy, but also
quantify the difference between these structures. Additionally, we
demonstrated the degree to which poor seeding structures can
significantly degrade performance of a competition structure, in some
cases making competition outcomes even worse than simply randomly
generating a set of ranks.

These results underscore the trade-off between tournament accuracy and
practical constraints such as time, cost, and entertainment value. While
round robin offers the highest fidelity to true rankings, it is often
impractical for large tournaments, highlighting the need for hybrid or
adaptive structures that balance accuracy with logistical feasibility.

Future work should extend this analysis to scenarios with more
tournament structures, varying numbers of competitors, incorporate
probabilistic strength models reflecting real-world uncertainty, and
explore alternative weighting schemes for different competitive
priorities (e.g., 4 salient ranks versus 8 salient ranking). Applying
these methods to actual tournament data could further validate their
usefulness for organizers aiming to design fair and informative
competitions.

# Acknowledgements {.unnumbered}

We thank the Department of Mathematics and Statistics at Loyola
University Chicago for their support and resources in conducting this
study. We thank Kailey Marie Lum for the suggestion of the term "salient
ranks". No external funding was received for this research.

# Supplementary Material {.unnumbered}

All supplementary material available at
<https://github.com/gjm112/tournaments>.

# References

<!-- author: |  -->

<!--   | \large Zach Culp \vspace{-1.1mm} -->

<!--   | \normalsize Department of Mathematics and Statistics \vspace{-1mm} -->

<!--   | \normalsize Loyola University Chicago \vspace{-1mm} -->

<!--   | \normalsize Chicago, IL 60660 \vspace{-1mm} -->

<!--   | -->

<!--   | \large Josie Peterburs \vspace{-1.1mm} -->

<!--   | \normalsize Department of Mathematics and Statistics \vspace{-1mm} -->

<!--   | \normalsize Loyola University Chicago \vspace{-1mm} -->

<!--   | \normalsize Chicago, IL 60660 \vspace{-1mm} -->

<!--   | -->

<!--   | \large Ryan McShane \vspace{-1.1mm} -->

<!--   | \normalsize University of Chicago \vspace{-1mm} -->

<!--   | \normalsize Chicago, IL 60660 \vspace{-1mm} -->

<!--   | -->

<!--   | \large Gregory J. Matthews \vspace{-1.1mm} -->

<!--   |    \normalsize  \vspace{-1mm} -->

<!--   | \normalsize Department of Mathematics and Statistics \vspace{-1mm} -->

<!--   | \normalsize Center for Data Science and Consulting \vspace{-1mm} -->

<!--   | \normalsize Loyola University Chicago \vspace{-1mm} -->

<!--   | \normalsize Chicago, IL 60660 \vspace{-1mm} -->

<!--   | \normalsize [`email`](mailto:ypu@something.edu) \vspace{-1mm} -->

<!--   | \normalsize [`zculp@luc.edu`](mailto:zculp@luc.edu) \vspace{-1mm} -->

<!--   | \normalsize [`jpeterburs@luc.edu`](mailto:jpeterburs@luc.edu) \vspace{-1mm} -->

<!--   | \normalsize [`rmcshane@uchicago.edu`](mailto:rmcshane@uchicago.edu) \vspace{-1mm} -->

<!--   | \normalsize [`gmatthews1@luc.edu`](mailto:gmatthews1@luc.edu) \vspace{-1mm} -->
